---
title: "Homework"
output: github_document
---

```{r}
library(tidyverse)
```

## Question 2
```{r}
mr_trash_wheel_df = 
  readxl::read_excel("data/202309_Trash_Wheel_Collection_Data.xlsx", sheet = 1, range = "A2:N586") |>
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30,
    trash_wheel_types = "Mr_trash_wheel"
  )
mr_trash_wheel_df$year = as.character(mr_trash_wheel_df$year)
```

# Import, clean, and organize the data for Professor Trash Wheel and Gwynnda

* Professor Trash Wheel

```{r}
prof_trash_wheel_df = 
  readxl::read_excel("data/202309_Trash_Wheel_Collection_Data.xlsx", sheet = 2, range = "A2:M108") |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30,
    trash_wheel_types = "professor_trash_wheel"
  )
prof_trash_wheel_df$year = as.character(prof_trash_wheel_df$year)
```


* Gwynnda Trash Wheel

```{r}
gwy_trash_wheel_df = 
  readxl::read_excel("data/202309_Trash_Wheel_Collection_Data.xlsx", sheet = 4, range = "A2:L157") |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons*500/30,
    trash_wheel_types = "Gwynnda_trash_wheel"
  )
gwy_trash_wheel_df$year = as.character(gwy_trash_wheel_df$year)
```


## Merging

```{r}
combine_sheet_df = 
  bind_rows(mr_trash_wheel_df,prof_trash_wheel_df,gwy_trash_wheel_df) |>
  janitor::clean_names() |> 
  relocate(trash_wheel_types)
```

* Calculate the total number of cigarette butts in 07/2021
```{r}
total_cigarettes = filter(gwy_trash_wheel_df, year == "2021", month == "July")
```


* Descriptions:
According to the results of combine sheet, we can see that there are **`r nrow(combine_sheet_df)` rows**, with **`r ncol(combine_sheet_df)` columns** in total. The key variables include **`r names(combine_sheet_df)`**. The total weight of trash collected by Professor Trash Wheel is **`r sum(prof_trash_wheel_df$weight_tons)`**. The total number of cigarette butts collected by Gwynnda in July of 2021 is **`r sum(total_cigarettes$cigarette_butts)`**

